---
title: "linkedin"
author: "nsren"
date: "2024-03-26"
output: html_document
---

##Set up enviroment
```{r}
library(tidyverse) #helps wrangle data
```

# STEP 1:
COLLECT DATA

# # Upload Divvy datasets (csv files) here

```{r}
q12020 <- read_csv("Divvy_Trips_2020_Q1.csv")
q42020 <- read_csv("Q4.csv")
q52020 <- read_csv("Q5.csv")
q62020 <- read_csv("Q6.csv")
q72020 <- read_csv("Q7.csv")
q82020 <- read_csv("Q8.csv")
q92020 <- read_csv("Q9.csv")
q102020 <- read_csv("Q10.csv")
q112020 <- read_csv("Q11.csv")
q122020 <- read_csv("Q12.csv")
```

# STEP 2: WRANGLE DATA AND COMBINE INTO A SINGLE FILE


# Convert start_station_id in q122020 to numaric so that they can stack correctly

```{r}
q122020 <- mutate(q122020,start_station_id  = as.numeric (start_station_id ))
```
## Convert end_station_id in q12020 and q122020 to numaric so that they can stack correctly
 
```{r}
q12020 <- mutate(q12020,end_station_id  = as.numeric (end_station_id ))
```

```{r}
q122020 <- mutate(q122020,end_station_id  = as.numeric (end_station_id ))
```

```{r}
is.numeric(q12020$end_station_id)

```
                  
# Stack individual quarter's data frames into one big data frame
```{r}
all_trips <- bind_rows(q12020, q42020,q52020,q62020,q72020,q82020,q92020,q102020,q112020,q122020)
```

# Remove lat, long, and end_station_id fields as we dont need
```{r}
all_trips <- all_trips %>%
  select(-c(start_lat, start_lng, end_lat, end_lng))
```
# STEP 3: CLEAN UP AND ADD DATA TO PREPARE FOR ANALYSIS

# Inspect the new table that has been created

```{r}
colnames(alltrips2) #List of column names
```

```{r}
summary(all_trips) #Statistical summary of data. Mainly for numerics
```

# There are a few problems in our data (2020) we will need to fix:
# (1) In the "member_casual" column, there are two names for members ("member" and
"Subscriber") and two names for casual riders ("Customer" and "casual"). We will need to
consolidate that from four to two labels.
# Begin by seeing how many observations fall under each usertype

```{r}
table(all_trips$member_casual)

```
```{r}
all_trips <- all_trips %>%
  mutate(member_casual = recode(member_casual
                                ,"Subscriber" = "member"
                                ,"Customer" = "casual"))
```
# Add columns that list the date, month, day, and day_of_week that provide additional
opportunities to aggregate the data.

```{r}
all_trips$date <- as.Date(all_trips$started_at) #The default format is yyyy-mm-dd
all_trips$month <- format(as.Date(all_trips$date), "%m")
all_trips$day <- format(as.Date(all_trips$date), "%d")
all_trips$day_of_week <- format(as.Date(all_trips$date), "%A")
```

# (3) # Add a "ride_length" calculation to all_trips (in seconds)
```{r}
all_trips$ride_length <- difftime(all_trips$ended_at,all_trips$started_at)
```
# Convert "ride_length" from Factor to numeric so we can run calculations on the data
```{r}
is.factor(all_trips$ride_length)
```
```{r}

all_trips$ride_length <- as.numeric(as.character(all_trips$ride_length))
is.numeric(all_trips$ride_length)
```
#Convert ride_length from from seconds into minutes
all_trips$ride_length <- as.numeric(all_trips$ride_length/60)


# Remove bad "negative" data
# We will create a new version of the dataframe (v2) since data is being removed
```{r}
all_trips <- all_trips[!(all_trips$start_station_name == "HQ QR" | all_trips$ride_length<0)]
```

# STEP 4:

CONDUCT DESCRIPTIVE ANALYSIS

# Descriptive analysis on ride_length (all figures in seconds)
```{r}
summary(all_trips_v2$ride_length)
```

# Compare members and casual users
```{r}
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = mean)
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = median)
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = max)
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = min)
```


# See the average ride time by each day for members vs casual users
```{r}

aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual + all_trips_v2$day_of_week,
          FUN = mean)
```
# Top 10 station

```{r}
 group_by(alltrips2,start_station_id,start_station_name)%>%

 summarize(Stations= n(),start_station_name)
```

# Notice that the days of the week are out of order. Let's fix that.
```{r}
all_trips$day_of_week <- ordered(all_trips$day_of_week, levels=c("Sunday", "Monday","Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
```


# Now, let's run the average ride time by each day for members vs casual users
```{r}
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual + all_trips_v2$day_of_week,
          FUN = mean)
```


# analyze ridership data by type and weekday
```{r}
all_trips_v2 %>%
  mutate(weekday = wday(started_at, label = TRUE)) %>% #creates weekday field using wday()
group_by(member_casual, weekday) %>% #groups by usertype and weekday
  summarise(number_of_rides = n() #calculates the number of rides and average duration
            ,average_duration = mean(ride_length)) %>% # calculates the average  duration
arrange(member_casual, weekday) 
```


